Competing Bandits:\\ The Perils of Exploration under Competition \thanks{This is an extended abstract of \citet{CompetingBandits-merged}, which in turn subsumes conference papers \citet{CompetingBandits-itcs18} and \citet{CompetingBandits-ec19}.}

### 2020-09-05 journal paper/content/related_work.tex
add bergemann valimaki paper comparisons

\xhdr{Exploration.} Multi-armed bandits (\emph{MAB}) is an elegant and tractable abstraction for tradeoff between \emph{exploration} and \emph{exploitation}: essentially, between acquisition and usage of information. MAB problems have been studied for many decades by researchers from computer science, operations research, statistics and economics, generating a vast and multi-threaded literature.  The most relevant thread concerns MAB with stochastic rewards and no auxiliary structure, which is the problem faced by each principal in our model (see Appendix~\ref{app:bg}). The basic model has been extended in many different directions, with a considerable amount of work on each: \eg payoffs with a specific structure (\eg combinatorial, linear, convex or Lipschitz), payoff distributions that change over time, and auxiliary payoff-relevant signals. This literature is covered in dedicated monographs,
\citep{Gittins-book11,Bubeck-survey12,slivkins-MABbook,LS19bandit-book}.%
\footnote{\citet{Gittins-book11} covers Markovian formulations, the other three are concerned with regret-minimization.}
Various connections to economics and game theory are detailed in 
books \citep{CesaBL-book,slivkins-MABbook} and surveys \citep{Bergemann-survey06,Horner-survey16}. Industrial applications are discussed in \citep{DS-arxiv}.

@@

\citet{keller2003price} studied the interplay of exploration and competition \emph{on prices}. This distinction leads to several important differences. First, the data externality -- that each customer brings a new data point to a given principal if and only if he chooses this principal -- is absolutely crucial to our setting, and absent in theirs. Second, the inputs to agents' decision rule are prices (which are directly controlled by the principals), rather than the quality of the chosen actions (which is not known a priori). Their focus is on learning the decision rule, whereas ours is on learning product quality. Third, they consider strategies that respond to competition and analyze Markov Perfect Equilibria, whereas we focus on the adoption of better bandit algorithms.

There is a superficial similarity --- in name only --- between this paper and the line of work on ``dueling bandits"
    \citep[\eg][]{Yue-dueling12,Yue-dueling-icml09}.
The latter is not about competing bandit algorithms, but rather about scenarios where in each round two arms are chosen to be presented to a user, and the algorithm only observes which arm has ``won the duel".

>>>>

\xhdr{Exploration.} Multi-armed bandits (\emph{MAB}) is an elegant and tractable abstraction for tradeoff between \emph{exploration} and \emph{exploitation}: essentially, between acquisition and usage of information. MAB problems have been studied for many decades by researchers from computer science, operations research, statistics and economics, generating a vast and multi-threaded literature.  The most relevant thread concerns MAB with stochastic rewards and no auxiliary structure, which is the problem faced by each principal in our model (see Appendix~\ref{app:bg}). The basic model has been extended in many different directions, with a considerable amount of work on each: \eg payoffs with a specific structure (\eg combinatorial, linear, convex or Lipschitz), payoff distributions that change over time, and auxiliary payoff-relevant signals.\footnote{There is a superficial similarity --- in name only --- between this paper and the line of work on ``dueling bandits" \citep[\eg][]{Yue-dueling12,Yue-dueling-icml09}.
The latter is not about competing bandit algorithms, but rather about scenarios where in each round two arms are chosen to be presented to a user, and the algorithm only observes which arm has ``won the duel".} This literature is covered in dedicated monographs,
\citep{Gittins-book11,Bubeck-survey12,slivkins-MABbook,LS19bandit-book}.%
\footnote{\citet{Gittins-book11} covers Markovian formulations, the other three are concerned with regret-minimization.}
Various connections to economics and game theory are detailed in 
books \citep{CesaBL-book,slivkins-MABbook} and surveys \citep{Bergemann-survey06,Horner-survey16}. Industrial applications are discussed in \citep{DS-arxiv}.

@@

\citet{keller2003price} studied the interplay of exploration and competition \emph{on prices}. This distinction leads to several important differences. First, the data externality -- that each customer brings a new data point to a given principal if and only if he chooses this principal -- is absolutely crucial to our setting, and absent in theirs. Second, the inputs to agents' decision rule are prices (which are directly controlled by the principals), rather than the quality of the chosen actions (which is not known a priori). Their focus is on learning the decision rule, whereas ours is on learning product quality. Third, they consider strategies that respond to competition and analyze Markov Perfect Equilibria (MPE), whereas we focus on the adoption of better bandit algorithms. \gaedit{\citet{bergemann1997market,bergemann2000experimentation} studied a model where there is an incumbent with known product quality and an entrant with unknown, but exogenously determined, quality. Similar to \citet{keller2003price}, the firms compete on prices and they characterize the MPE of dynamic competition between the two firms. In their model, consumer purchases of the entrant's product result in an informational externality that allows both the firms and consumers to gather information on the entrant's product quality. However, we focus on a model where firms' product quality is endogenous and the focus of exploration in a context where the price mechanism is absent.}
